Open Data, Open Source and Open Standards in chemistry - The Blue Obelisk five years on

### 2011-05-13 paper.tex
Added subsection about computational chemistry and info/cite to cclib.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Blue Obelisk Open Source software projects}
    (Description if necessary XXXXXXXXXXXXXXX. Add citations to project names.)
    \par \mbox{}
    \par
    \mbox{
      \begin{tabular}{|c|c|c|}
        %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
        \hline Name & Website & Description? or Lead Developer? \\ \hline
        \multicolumn{3}{|c|}{Cheminformatics toolkits} \\ \hline
        Chemistry Development Kit (CDK) \cite{Steinbeck2003, Steinbeck2006} & http://cdk.sf.net & XXXX  \\ \hline
        Cinfony & http://cinfony.googlecode.com & Noel O'Boyle, Python interface to toolkits \\ \hline
        Indigo & http://ggasoftware.com/opensource/indigo & GGA Software \\ \hline
        Open Babel & http://openbabel.org & Geoffrey Hutchison et al \\ \hline
        RDKit & http://rdkit.org & Greg Landrum \\ \hline
        ChemDoodle Web Components & http://web.chemdoodle.com & iChemLabs \\ \hline
        \multicolumn{3}{|c|}{Integration} \\ \hline
        CDK-Taverna \cite{Kuhn:2010p4001} & http://cdk-taverna.blah.XXX & Christoph Steinbeck, Workflow \\ \hline
        A3 & ..  & .  \\ \hline
        \multicolumn{3}{|c|}{Interconversion} \\ \hline
        OSRA & http://osra.sf.net & Igor Filippov, Image to structure \\ \hline
        OPSIN & http://opsin.ch.cam.ac.uk & Daniel Lowe, Name to structure \\ \hline
        A3 & ..  & .  \\ \hline
        \multicolumn{3}{|c|}{Structure Databases} \\ \hline
        OrChem \cite{RijnbeekS10} & http://orchem.sourceforge.net & Christoph Steinbeck, Oracle-based Chemical Database Engine \\ \hline
        MyChem  & http://XXX.sourceforge.net & Jerome Pansanel, MySQL-based Chemical Database Engine \\ \hline
        PGChem  & http://XXX.sourceforge.net & Ernst-Georg Schmid, PostgreSQL-based Chemical Database Engine \\ \hline
        Bingo  & http://XXX.XXX.XXX & Dmitry Pavlov, Oracle-based Chemical Database Engine \\ \hline

      \end{tabular}
      }
  \subsection*{Table 2 - Open Data in chemistry.}
    Overview of major open chemical data available under a license or waiver
    compatible with the Panton Principles.
    \par \mbox{}
    \par
    \mbox{
\begin{tabular}{|l|l|l|}
  %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
  \hline Name & License/Waiver & Description \\ \hline
  ChemPedia & CC0 & Crowd-sourced chemical names. Project discontinued. \\ \hline
  CrystalEye     & PPDL & Crystal structures from primary literature. \\ \hline
  ONS Solubility & CC0 & Solubility data for various solvents. \\ \hline
\end{tabular}
      }

>>>>

\subsection*{Facilitating chemical computations}

An area where cheminformatic tools can make a signficant impact is supporting
chemical computations and interpreting their results.
This is especially true for electronic structure calculations since many
quantum chemistry methods already have a long tradition in the scientific
community -- most prominently the Hartree-Fock method and density functional theory,
as well as more accurate variations of these.
Many methods have been implemented and reimplemented in a variety of software
packages over the last 30 years.
When a computational chemist performs calculations using several programs,
he/she will find that the format of their input files and outputs
differ enourmously.
Sometimes even different version of the same program implement different input
syntax and output formats.
Furthermore, some of these programs provide additional algorithms such as
populations analyses, which in general are applicable to all packages.

Obtaining electronic structure is often the first step in a research pipeline,
after which data can be further analyzed or visualized.
Preparing inputs and some basic analyses, usually with the
goal of visualization, are often a part of graphical molecular editors.
Avogadro for example is capable of generating input files for several
electronic structure packages, and visualizes molecular orbitals.

The Blue Obelisk project  cclib\cite{cclib} is dedicated to parsing output
from computational chemistry programs.
It's main goal is to read as many different kinds of results from as
many different packages as possible.
After parsing, cclib provides all results in standardized form as attributes
of a python object Python.
This means that if a developer implements an algorithm based on such an API,
they can feed it output files from any of the supported computational software.
The current version of cclib (1.0.1) supports 8 different computational chemistry
codes and extracts over 30 different calculations results.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Blue Obelisk Open Source software projects}
    (Description if necessary XXXXXXXXXXXXXXX. Add citations to project names.)
    \par \mbox{}
    \par
    \mbox{
      \begin{tabular}{|c|c|c|}
        %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
        \hline Name & Website & Description? or Lead Developer? \\ \hline
        \multicolumn{3}{|c|}{Cheminformatics toolkits} \\ \hline
        Chemistry Development Kit (CDK) \cite{Steinbeck2003, Steinbeck2006} & http://cdk.sf.net & XXXX  \\ \hline
        Cinfony & http://cinfony.googlecode.com & Noel O'Boyle, Python interface to toolkits \\ \hline
        Indigo & http://ggasoftware.com/opensource/indigo & GGA Software \\ \hline
        Open Babel & http://openbabel.org & Geoffrey Hutchison et al \\ \hline
        RDKit & http://rdkit.org & Greg Landrum \\ \hline
        ChemDoodle Web Components & http://web.chemdoodle.com & iChemLabs \\ \hline
        \multicolumn{3}{|c|}{Integration} \\ \hline
        CDK-Taverna \cite{Kuhn:2010p4001} & http://cdk-taverna.blah.XXX & Christoph Steinbeck, Workflow \\ \hline
        A3 & ..  & .  \\ \hline
        \multicolumn{3}{|c|}{Interconversion} \\ \hline
        OSRA & http://osra.sf.net & Igor Filippov, Image to structure \\ \hline
        OPSIN & http://opsin.ch.cam.ac.uk & Daniel Lowe, Name to structure \\ \hline
        A3 & ..  & .  \\ \hline
        \multicolumn{3}{|c|}{Structure Databases} \\ \hline
        OrChem \cite{RijnbeekS10} & http://orchem.sourceforge.net & Christoph Steinbeck, Oracle-based Chemical Database Engine \\ \hline
        MyChem  & http://XXX.sourceforge.net & Jerome Pansanel, MySQL-based Chemical Database Engine \\ \hline
        PGChem  & http://XXX.sourceforge.net & Ernst-Georg Schmid, PostgreSQL-based Chemical Database Engine \\ \hline
        Bingo  & http://XXX.XXX.XXX & Dmitry Pavlov, Oracle-based Chemical Database Engine \\ \hline
        cclib  & http://cclib.sourceforge.net & Python library for parsing and interpreting computational chemistry results \\ \hline
        \end{tabular}
      }
  \subsection*{Table 2 - Open Data in chemistry.}
    Overview of major open chemical data available under a license or waiver
    compatible with the Panton Principles.
    \par \mbox{}
    \par
    \mbox{
\begin{tabular}{|l|l|l|}
  %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
  \hline Name & License/Waiver & Description \\ \hline
  ChemPedia & CC0 & Crowd-sourced chemical names. Project discontinued. \\ \hline
  CrystalEye     & PPDL & Crystal structures from primary literature. \\ \hline
  ONS Solubility & CC0 & Solubility data for various solvents. \\ \hline
\end{tabular}
      }
